SIR Model With Testing Flows and Isolation

### 2021-01-18 codes/sir_macpan.Rnw
forcast updated
updated with BMB comments

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Step 1, SIR and MacPan}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

<<Read Data from Mike's repo, echo=T>>=
library(McMasterPandemic)
library(readr)
library(dplyr)
library(ggplot2); theme_set(theme_bw())
library(directlabels)

covid_url <- "https://raw.githubusercontent.com/wzmli/COVID19-Canada/master/git_push/clean.Rout.csv"
df_covid <- read_csv(covid_url)

## BMB: the fit was using *cumulative* cases before, 
## Ali I used confirmed_positive before
plot_pos_prov <- ggplot(df_covid,aes(x=Date,y=newConfirmations,col=Province))+
  geom_point()

direct.label(plot_pos_prov,"last.bumpup")

# Assume we are just looking at Ontario's epidemic case:

startdate <- as.Date("2020-12-11")
enddate <- as.Date("2021-01-14")
## BMB: if we use the original, short time series
##   Error in stats::filter(i, kern, sides = 1) : 
##     'filter' is longer than time series

df_covid_on <- (df_covid 
                %>% filter(Province =="ON", Date >= startdate) # Ontario
                %>% mutate(value= round(newConfirmations), var= "report", date=Date)
                %>% select(date, value, var)
                %>% na.omit()
)
## BMB: check length, beginning
nrow(df_covid_on)
head(df_covid_on,1)

## beta0 (and E0) the only parameter we're going to optimize:
params <- read_params("ICU1.csv")
## adjust to ON pop size, approx starting value:
## about 1800 cases reported
## adjust ascertainment rate (fraction of cases reported) to 1/3
## (was 0.1, which was reasonable at beginning of epidemic)
## (will just rescale results)
## so 1800 cases -> prevalence of approx 3*1800
## initial conditions automatically spread E0 over the eigenvector
params <- update(params, N=15e6,E0=3*1800, c_prop=1/3)

@@

params1obs <- update(params, obs_disp=20)

@@

## fit beta0 based on the report data:
fitted.mod <- calibrate(
  data = df_covid_on
  , start_date = startdate
  ## skip breaks that are present by default:
, time_args = list(break_dates = NULL)
  , base_params = params1obs
, opt_pars = opt_pars
 , debug_plot = TRUE # instructive plotting during optimization
)

@@

sim0 <- run_sim(pp,state,start_date=startdate,end_date=enddate)
gg0 <- (ggplot(sim0,aes(x=date))
	+ geom_point(aes(y=report))
)
print(gg0)

## Question: the runsim gives reported cases starrting from 2465 cases and grows, it doesn't match the reported data! I think I am missing some scaling here?

@@

time_pars <- data.frame(Date=as.Date(startdate:enddate)
	, Symbol="beta0"
	, Relative_value= pp[["beta0"]] ## Here I am using the caliberated Beta0
)
	# , stringsAsFactors=FALSE)

time_pars2 <- 
    data.frame(Date=as.Date(startdate:enddate)
             , Symbol="beta0"
             , Relative_value = 
                   c(rep(pp[["beta0"]], length(startdate:lockdown)-1)
                   , seq(pp[["beta0"]],pp[["beta0"]]-0.1,length.out = length(lockdown:enddate))
                     )
               )
print(time_pars2)

@@

print(gg0
		+ geom_point(data=sim0_t_reduce, aes(x=date,y=incidence), color="red")
)

>>>>

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Step 1 SIR and MacPan}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

<<Read Data from Mike's repo, echo=T>>=
library(McMasterPandemic)
library(readr)
library(dplyr)
library(ggplot2); theme_set(theme_bw())
library(directlabels)
path <- "/home/ag/projects/SIRmodel/codes/"
covid_url <- "https://raw.githubusercontent.com/wzmli/COVID19-Canada/master/git_push/clean.Rout.csv"
covid_can <- read_csv(covid_url)
fn_covid_can = paste0(path,"covid_can.csv")
write_csv(covid_can,fn_covid_can)

## Ali I used confirmed_positive before
## BMB: the fit was using *cumulative* cases before, 
plot_pos_prov <- ggplot(covid_can,
                        aes(x=Date,y=newConfirmations,col=Province))+
                  geom_point()
direct.label(plot_pos_prov,"last.bumpup")

# Assume we are just looking at Ontario's epidemic case:
startdate <- as.Date("2020-12-11")
enddate <- as.Date("2021-01-14")
## BMB: if we use the original, short time series
##   Error in stats::filter(i, kern, sides = 1) : 
##     'filter' is longer than time series

df_covid_on <- (covid_can 
                %>% filter(Province =="ON", Date >= startdate) # Ontario
                %>% mutate(value= round(newConfirmations), var= "report", date=Date)
                %>% select(date, value, var)
                %>% na.omit()
)
## BMB: check length, beginning
nrow(df_covid_on)
head(df_covid_on,1)
tail(df_covid_on)

## beta0 (and E0) the only parameter we're going to optimize:
params <- read_params("ICU1.csv")
## adjust to ON pop size, approx starting value:
## about 1800 cases reported
## adjust ascertainment rate (fraction of cases reported) to 1/3
## (was 0.1, which was reasonable at beginning of epidemic)
## (will just rescale results)
## so 1800 cases -> prevalence of approx 3*1800
## initial conditions automatically spread E0 over the eigenvector
params <- update(params, N=15e6,E0=exp(11.2), c_prop=1/3, beta=-1.5)

@@

params1obs <- update(params, obs_disp=20) #set dispersion params to 20

@@

## fit beta0 based on the report data:
fitted.mod <- calibrate(
  data = df_covid_on
  , start_date = startdate
  ## skip breaks that are present by default:
  , time_args = list(break_dates = NULL)
  , base_params = params1obs
  , opt_pars = opt_pars
  , debug_plot = TRUE # instructive plotting during optimization
  , use_DE= FALSE
  , DE_lwr=c(6,-2,-2),
  , DE_upr=c(12,2,2)
)

@@

sim0 <- run_sim(pp,state,start_date=startdate,end_date=enddate)
gg0 <- (ggplot(sim0,aes(x=date))
    	+ geom_point(aes(y=report))
        )
print(gg0)
# Ben
print(gg0 + geom_point(data=df_covid_on,aes(y=value),colour="red"
                       ))

@@

time_pars <- data.frame(Date=as.Date(startdate:enddate)
	, Symbol="beta0"
	, Relative_value= 1 ## Here I am using the caliberated Beta0
)

time_pars2 <- 
    data.frame(Date=seq(lockdown,enddate,by=1)
             , Symbol="beta0"
             , Relative_value = 
               seq(1,.5,length=enddate-lockdown+1) )

print(time_pars2)

@@

print(gg0
		+ geom_point(data=sim0_t_reduce, aes(x=date,y=report), color="blue")
		+ geom_point(data=df_covid_on,aes(y=value),colour="red"  )
      )
## the blue (black) curve is forecast with (without) lockdown impact
## the red curve is actual data

## Question: How to add confidence Int?
